Faithful Variable Screening for High-Dimensional Convex Regression

### 2014-07-11 tex/scam/appendix.tex
some changes

Define $\mathcal{G}$ as $\{ f - f^* + \bar{f}^* \,:\, f \in \mathcal{C}^s_L\}$ as the set of convex additive functions centered around the function $f^* - \bar{f}^*$. Then, $\| h \|_n \leq \| h \|_\infty \leq 4sLb$ for all $h \in \mathcal{G}$.\\

According to theorem~\ref{thm:chaining}, for all $\epsilon > \frac{1}{\sqrt{n}}\sigma c \int_0^R \sqrt{ \log N_2(t, \mathcal{G}) }dt \vee R$,
\[
P\Big( \sup_{h \in \mathcal{G}} \langle w, h \rangle_n \geq \epsilon \Big) \leq
  4 \exp \Big( - \frac{ n \epsilon^2}{ c R^2 \sigma^2} \Big)
\]
where $R = 4sLb$ for our purpose.

Restated, we have that, with probablity at least $1-\delta$,
\[
\sup_{h \in \mathcal{G}} | \langle w, h \rangle_n | \leq 
   c R \sigma \sqrt{ \frac{1}{n} \log \frac{4}{\delta}} + 
      \Big( \int_0^R \sqrt{\log N_2(t, \mathcal{G})}dt \vee R \Big)\, 
       c \sigma \sqrt{\frac{1}{n}} 
\]

Now we evaluate the integral. Since $N_{\|\cdot\|_n}(t, \mathcal{G}) \leq N_\infty(t, \mathcal{G})$, we know that $\sqrt{\log N_{\|\cdot\|_n}(t, \mathcal{G})} \leq \sqrt{C s^{1.5} b L} t^{-1/4}$.

\begin{align*}
\int_0^R \sqrt{\log N_{\|\cdot\|_n}(t, \mathcal{G})} dt &\leq 
      \sqrt{C s^{1.5} b L} \int_0^R t^{-1/4} dt \\ 
 &= \sqrt{C s^{1.5} b L} \, \frac{4}{3} R^{3/4} \\
 &= \sqrt{C s^{1.5} b L} \, c (sLb)^{3/4} \\
 &\leq c (s b L)^2
\end{align*}

Coming back, we have, with probability at least $1-\delta$,
\begin{align*}
\sup_{h \in \mathcal{G}} | \langle w, h \rangle | &\leq 
   c sLb \sigma \sqrt{ \frac{1}{n} \log \frac{4}{\delta} } + 
    c (sLb)^2 \sigma \sqrt{ \frac{1}{n} } \\
 &\leq c (sLb)^2\sigma \sqrt{ \frac{1}{n} \log \frac{4}{\delta} }
\end{align*}

@@

For a $h \in \mathcal{G}$, let $h_\epsilon$ denote a function in the $\epsilon$-cover of $\mathcal{G}$ closest to $h$. It obviously must be that $\| h - h_\epsilon \|_n \leq \| h - h_\epsilon \|_\infty \leq \epsilon$.\\

@@

\subsubsection{Covering Number for Lipschitz Convex Functions}

\begin{definition}
$\{ f_1,..., f_N\} \subset \mathcal{C}[b,B,L]$ is an $\epsilon$-covering of $\mathcal{C}[b,B,L]$ if for all $f \in \mathcal{C}[b,B,L]$, there exist $f_i$ such that $\| f - f_i \|_\infty \leq \epsilon$.

We define $N_\infty( \epsilon, \mathcal{C}[b,B,L])$ as the size of the minimum covering.
\end{definition}

\begin{lemma} (Bronshtein 1974)
\[
\log N_\infty (\epsilon, \mathcal{C}[b,B,L]) \leq C\left( \frac{bBL}{\epsilon} \right)^{1/2}
\]
For some absolute constant $C$.
\end{lemma}

\begin{lemma}
\[
\log N_\infty( \epsilon, \mathcal{C}^s[b,B,L])  \leq C s \left(\frac{bBLs}{\epsilon}\right)^{1/2}
\]
For some absolute constant $C$.
\end{lemma}

\begin{proof}
Let $f = \sum_{k=1}^s f_k$ be a convex additive function. Let $\{ f'_k \}_{k=1,..,s}$ be $k$ functions from a $\frac{\epsilon}{s}$ $L_\infty$ covering of $\mathcal{C}[b,B,L]$.

Let $f' \coloneqq \sum_{k=1}^s f'_k$, then 
\[
\| f' - f \|_{\infty} \leq \sum_{k=1}^s \| f_k - f'_k \|_\infty \leq s \frac{\epsilon}{s} \leq \epsilon
\]

Therefore, a product of $s$ $\frac{\epsilon}{s}$-coverings of univariate functions induces an $\epsilon$-covering of the additive functions.
\end{proof}

In our proofs, we will use the Dudley's chaining: [TODO:cite van de geer]

\begin{theorem} (Dudley's Chaining) \\
\label{thm:chaining}
Let $\mathcal{G} = \{ g : \| g \|_n \leq R \}$. Let $M( \epsilon, R)$ be the size of the minimal $\epsilon$-covering of $\mathcal{G}$ with respect to the $\| \cdot \|_n$ norm. Suppose $w = (w_1, ..., w_n)$ is a vector of i.i.d. subgaussian random variables with scale $\sigma$.\\

Suppose $\delta > 0$ is such that
\[
\sqrt{n} \delta \geq \sigma \Big( 
   14 \sum_{s=0}^\infty 2^{-s} \sqrt{ \log M( 2^{-s} R, \mathcal{G})} 
      \Big)\vee 70 \log 2 R
\]
Then we have that 
\[
P\Big( \sup_{g \in \mathcal{G}} \langle w, g \rangle_n \geq \delta \Big) \leq
  4 \exp \Big( - \frac{n \delta^2}{(70R)^2\sigma^2} \Big)
\]
\end{theorem}

For convenience, we can upper bound the metric-entropy sum with an integral: \\
$\ds \sum_{s=0}^\infty 2^{-s} \sqrt{ \log M( 2^{-s} R, \mathcal{G})} 
  \leq \int_0^R \sqrt{\log M(t, \mathcal{G}) } dt
$

>>>>

Define $\mathcal{G}$ as $\{ f - f^* + \bar{f}^* \,:\, f \in \mathcal{C}^s_L\}$ as the set of convex additive functions centered around the function $f^* - \bar{f}^*$.

% Then, $\| h \|_n \leq \| h \|_\infty \leq 4sLb$ for all $h \in \mathcal{G}$.\\

% According to theorem~\ref{thm:chaining}, for all $\epsilon > \frac{1}{\sqrt{n}}\sigma c \int_0^R \sqrt{ \log N_2(t, \mathcal{G}) }dt \vee R$,
% \[
% P\Big( \sup_{h \in \mathcal{G}} \langle w, h \rangle_n \geq \epsilon \Big) \leq
%   4 \exp \Big( - \frac{ n \epsilon^2}{ c R^2 \sigma^2} \Big)
% \]
% where $R = 4sLb$ for our purpose.

% Restated, we have that, with probablity at least $1-\delta$,
% \[
% \sup_{h \in \mathcal{G}} | \langle w, h \rangle_n | \leq 
%    c R \sigma \sqrt{ \frac{1}{n} \log \frac{4}{\delta}} + 
%       \Big( \int_0^R \sqrt{\log N_2(t, \mathcal{G})}dt \vee R \Big)\, 
%        c \sigma \sqrt{\frac{1}{n}} 
% \]

% Now we evaluate the integral. Since $N_{\|\cdot\|_n}(t, \mathcal{G}) \leq N_\infty(t, \mathcal{G})$, we know that $\sqrt{\log N_{\|\cdot\|_n}(t, \mathcal{G})} \leq \sqrt{C s^{1.5} b L} t^{-1/4}$.

% \begin{align*}
% \int_0^R \sqrt{\log N_{\|\cdot\|_n}(t, \mathcal{G})} dt &\leq 
%       \sqrt{C s^{1.5} b L} \int_0^R t^{-1/4} dt \\ 
%  &= \sqrt{C s^{1.5} b L} \, \frac{4}{3} R^{3/4} \\
%  &= \sqrt{C s^{1.5} b L} \, c (sLb)^{3/4} \\
%  &\leq c (s b L)^2
% \end{align*}

% Coming back, we have, with probability at least $1-\delta$,
% \begin{align*}
% \sup_{h \in \mathcal{G}} | \langle w, h \rangle | &\leq 
%    c sLb \sigma \sqrt{ \frac{1}{n} \log \frac{4}{\delta} } + 
%     c (sLb)^2 \sigma \sqrt{ \frac{1}{n} } \\
%  &\leq c (sLb)^2\sigma \sqrt{ \frac{1}{n} \log \frac{4}{\delta} }
% \end{align*}

@@

By Corollary~\ref{cor:convexbracket_ln}, there exists a bracketing $\mathcal{G}_\epsilon$ of $\mathcal{G}$ which has sized bounded by [TODO:...], which is an $\epsilon$-bracketing with respect to the $L_2(P)$ norm, and which is also an $(\epsilon + \epsilon_{n,\delta})$-bracketing with respect to the $L_n$ norm with probability at least $1-\delta$.

For a $h \in \mathcal{G}$, let $h_\epsilon$ denote a function in the $\epsilon$-bracketing of $\mathcal{G}$ closest to $h$. It obviously must be that $\| h - h_\epsilon \|_n \leq \| h - h_\epsilon \|_\infty \leq \epsilon$.\\

@@

\subsubsection{Bracketing Number for Convex Functions}

\begin{definition}
% upper, lower, epsilon room
% metric
% function space
Let $\mathcal{C}$ be a set of functions. For a given $\epsilon$ and metric $\rho$ (which we take to be $L_2$ or $L_2(P)$), we define a \textbf{bracketing} of $\mathcal{C}$ to be a set of pairs of functions $\{ (f^L, f^U) \}$ satisfying (1) $\rho( f^L, f^U) \leq \epsilon$ and (2) for any $f \in \mathcal{C}$, there exist a pair $(f^L, f^U)$ where $f^U \geq f \geq f^L$.

We let $N_{[]}(\epsilon, \mathbf{C}, \rho)$ denote the size of the smallest bracketing of $\mathcal{C}$
\end{definition}

\begin{proposition} (Proposition 16 in \cite{kim2014global})\\
\label{prop:convexbracket}
Let $\mathcal{C}$ be the set of convex functions supported on $[-b, b]^d$ and uniformly bounded by $B$. Then there exist constants $\epsilon_3$ and $K^{**}$, dependent on $d$, such that
\[
\log N_{[]} (2\epsilon, \mathcal{C}, L_2) \leq K^{**} \left( \frac{2bB}{\epsilon} \right)^{d/2}
\]
for all $\epsilon \in (0, B \epsilon_3]$.
\end{proposition}

It is trivial to extend Kim and Samworth's result to $L_2(P)$ norm for a distribution $P$ with an uniformly upper bounded density.

\begin{corollary}
\label{cor:convexbracket_lp}
Let $P$ be a distribution with a density $p$ upper bounded by $c_u$. Let $\mathcal{C}, b, B, \epsilon_3, K^{**}$ be defined as in Proposition~\ref{prop:convexbracket}. Then,
\[
\log N_{[]} (2\epsilon, \mathcal{C}, L_2(P)) \leq K^{**} \left( \frac{2bB c_u}{\epsilon} \right)^{d/2}
\]
\end{corollary}

Given samples $X_1,..., X_n \in \R^p$, we define $L_n$-norm as $\| f \|^2_n \equiv \frac{1}{n}\sum_{i=1}^n f(X_i)^2$. The following result follows from Corollary~\ref{cor:convexbracket_lp} directly by an union bound.

\begin{corollary}
\label{cor:convexbracket_ln}
Let $X_1,...,X_n$ be random samples from a distribution $P$. Let $1 > \delta > 0$. Let $\mathcal{C}_\epsilon$ be an $\epsilon$-bracketing of $\mathcal{C}$ with respect to the $L_2(P)$-norm whose size is at most $N_{[]}( 2\epsilon, \mathcal{C}, L_2(P))$. \\

Then, with probability at least $1-\delta$, for all pairs $(f^L, f^U) \in \mathcal{C}_\epsilon$, $\frac{1}{n} \sum_{i=1}^n ( f^L(X_i) - f^U(X_i)^2 \leq \epsilon + \epsilon_{n, \delta}$, where 
$\epsilon_{n,\delta} \equiv 
B^2 \sqrt{ \frac{ \log N_{[]}(2\epsilon, \mathcal{C}, L_2(P)) + \log \frac{1}{\delta}}{2n}} = 
B^2 \sqrt{ \frac{K^{**} (2bBc_u)^{d/2} + \log \frac{1}{\delta}}{2\epsilon^{d/2} n} }$
\end{corollary}

\begin{proof}
$(f^L(X_i) - f^U(X_i))^2$ is at most $B^2$. There are $N_{[]}(2\epsilon, \mathcal{C}, L_2(P))$ pairs $(f^L, f^U)$. The corollary follows from a direct application of union bound and Hoeffding Inequality.
\end{proof}

% \subsubsection{Covering Number for Lipschitz Convex Functions}

% \begin{definition}
% $\{ f_1,..., f_N\} \subset \mathcal{C}[b,B,L]$ is an $\epsilon$-covering of $\mathcal{C}[b,B,L]$ if for all $f \in \mathcal{C}[b,B,L]$, there exist $f_i$ such that $\| f - f_i \|_\infty \leq \epsilon$.

% We define $N_\infty( \epsilon, \mathcal{C}[b,B,L])$ as the size of the minimum covering.
% \end{definition}

% \begin{lemma} (Bronshtein 1974)
% \[
% \log N_\infty (\epsilon, \mathcal{C}[b,B,L]) \leq C\left( \frac{bBL}{\epsilon} \right)^{1/2}
% \]
% For some absolute constant $C$.
% \end{lemma}

% \begin{lemma}
% \[
% \log N_\infty( \epsilon, \mathcal{C}^s[b,B,L])  \leq C s \left(\frac{bBLs}{\epsilon}\right)^{1/2}
% \]
% For some absolute constant $C$.
% \end{lemma}

% \begin{proof}
% Let $f = \sum_{k=1}^s f_k$ be a convex additive function. Let $\{ f'_k \}_{k=1,..,s}$ be $k$ functions from a $\frac{\epsilon}{s}$ $L_\infty$ covering of $\mathcal{C}[b,B,L]$.

% Let $f' \coloneqq \sum_{k=1}^s f'_k$, then 
% \[
% \| f' - f \|_{\infty} \leq \sum_{k=1}^s \| f_k - f'_k \|_\infty \leq s \frac{\epsilon}{s} \leq \epsilon
% \]

% Therefore, a product of $s$ $\frac{\epsilon}{s}$-coverings of univariate functions induces an $\epsilon$-covering of the additive functions.
% \end{proof}

% In our proofs, we will use the Dudley's chaining: [TODO:cite van de geer]

% \begin{theorem} (Dudley's Chaining) \\
% \label{thm:chaining}
% Let $\mathcal{G} = \{ g : \| g \|_n \leq R \}$. Let $M( \epsilon, R)$ be the size of the minimal $\epsilon$-covering of $\mathcal{G}$ with respect to the $\| \cdot \|_n$ norm. Suppose $w = (w_1, ..., w_n)$ is a vector of i.i.d. subgaussian random variables with scale $\sigma$.\\

% Suppose $\delta > 0$ is such that
% \[
% \sqrt{n} \delta \geq \sigma \Big( 
%    14 \sum_{s=0}^\infty 2^{-s} \sqrt{ \log M( 2^{-s} R, \mathcal{G})} 
%       \Big)\vee 70 \log 2 R
% \]
% Then we have that 
% \[
% P\Big( \sup_{g \in \mathcal{G}} \langle w, g \rangle_n \geq \delta \Big) \leq
%   4 \exp \Big( - \frac{n \delta^2}{(70R)^2\sigma^2} \Big)
% \]
% \end{theorem}

% For convenience, we can upper bound the metric-entropy sum with an integral: \\
% $\ds \sum_{s=0}^\infty 2^{-s} \sqrt{ \log M( 2^{-s} R, \mathcal{G})} 
%   \leq \int_0^R \sqrt{\log M(t, \mathcal{G}) } dt
% $

### 2014-07-11 tex/scam/introduction.tex
some changes

Shape restrictions such as monotonicity, convexity, and concavity
provide a natural way of limiting the complexity of many statistical
estimation problem.  Shape-constrained estimation is not as well
understood as more traditional nonparametric estimation involving
smoothness constraints.  For instance, as of this writing the minimax
rate of convergence for multivariate convex regression has yet to be
rigorously established in full generality.  Even the one-dimensional
case is challenging, and has been of recent interest
\citep{guntusen:13}.

In this paper we study the problem of variable selection
in convex regression.   Assuming that the regression
function is convex and sparse, in order to select the
relevant variables we 
show that it suffices to estimate a sum of $p$ one-dimensional convex
functions, leading to significant computational and statistical
advantages .  In addition, we introduce algorithms and supporting
statistical theory for a practical, effective approach to this
variable selection problem.

>>>>

We consider the problem of estimating a convex function of several
variables from noisy values of the function at a finite sample of
input points.

Recent work \cite{Guntu:12,Guntu:13} shows that the
minimax rate for convex function estimation in $p$ dimensions is
$n^{-4/(4+p)}$.  Loosely speaking, this shows that the
geometric convexity constraint is statistically equivalent to
requiring two derivatives of the function, and thus is
subject to the same curse of dimensionality.

However, if the function is sparse, with $s\ll p$ relevant variables,
then the faster rate $n^{-4/(4+s)}$ may be achievable if the $s$
variables can be identified.  To determine the relevant variables, we
show that it suffices to estimate a sum of $p$ one-dimensional convex
functions, leading to significant computational and statistical
advantages.  In addition, we introduce algorithms and supporting
statistical theory for a practical, effective approach to this
variable selection problem.
